# Metabolomic Pathway Enrich Analysis

During the master a software called [MetaboAnalyst](https://www.metaboanalyst.ca/MetaboAnalyst/ModuleView.xhtml) was teached and allows the user to do a **metabolite enrichment for KEGG pathways**. An enrichment for activated and repressed metabolites was performed using **Chlorella variabilis** metabolic pathways as an orthologue of *Klebsormidium nitens*. As there were few metabolites Fisher's test was employed and to know the importance of each node the betweeness metric was used.


```{r echo=F,warning=F}
# Necessary packages:
library(ggplot2)
library(MetBrewer)
library(cowplot)
library(plotly)
library(ggrepel)


# Enrichment for activated metabolites:
pathway_results_act_chlo <- read.csv("Download_chlo_act/pathway_results.csv",header = T,sep = ",")
# head(pathway_results_act_chlo)
pathways_act_chlo <- pathway_results_act_chlo$X
colnames(pathway_results_act_chlo)[1] <- "KEGG_Pathways"
colnames(pathway_results_act_chlo)[6] <- "log10_p.value"
rownames(pathway_results_act_chlo) <- pathways_act_chlo

# Enrichment for repressed metabolites:
pathway_results_rep_chlo <- read.csv("Download_rep_chlo/pathway_results.csv",header = T,sep = ",")
# head(pathway_results_rep_chlo)
pathways_rep_chlo <- pathway_results_rep_chlo$X
colnames(pathway_results_rep_chlo)[1] <- "KEGG_Pathways"
colnames(pathway_results_rep_chlo)[6] <- "log10_p.value"
rownames(pathway_results_rep_chlo) <- pathways_rep_chlo

```

As a result it generates two tables: one with the conversions that the program needs to do called [name_map.csv](Download/name_map.csv) and another one with the data of the KEGG path enrichment called [pathway_results.csv](Download/pathway_results.csv). To visualize it, bubble plots were made showing the pvalue of each enriched pathway versus the impact exerted by that metabolite on that pathway.


```{r warning=F,echo=F}

# Plots title:

a <- list(
  text = "Enriquecimiento Metabolitos Activados CHLO", xref = "paper",yref = "paper",yanchor = "bottom", xanchor = "center",align = "center",x = 0.5,y = 1, showarrow = FALSE)

b <- list(
  text = "Enriquecimiento Metabolitos Reprimidos CHLO",xref = "paper",yref = "paper",yanchor = "bottom", xanchor = "center",align = "center",x = 0.5,y = 1, showarrow = FALSE)


# Activated
Activados_chlo <- ggplot(pathway_results_act_chlo, aes(x=Impact, y=log10_p.value, size = Hits, color=log10_p.value,
                                         text= paste0("</br> KEGG: " ,KEGG_Pathways))) +
  geom_point(alpha=0.7)+
  geom_label_repel(aes(label = KEGG_Pathways),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  size=5.5) +
  scale_color_gradientn(colors = met.brewer("Gauguin",type = "continuous")) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "white"),
        panel.grid.minor = element_line(colour = "white"),
        axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left   = element_line(color = 'black'),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 15),
        axis.text=element_text(size=18),
        axis.title=element_text(size=18,face="bold"))+
  ylab("-log10(p.value)") +
  xlab("Pathway impact value")


ggsave2("Activados_met.png",width = 8,height = 8)


Activados_plotly_chlo <- ggplotly(Activados_chlo,x = ~Impact, y = ~log10_p.value,
                             tooltip = "text",
                             width = 1100, height = 900)%>%
               layout(annotations = a)

# Repressed

Reprimidos_chlo <- ggplot(pathway_results_rep_chlo, aes(x=Impact, y=log10_p.value, size = Hits, color=log10_p.value,
                                              text= paste0("</br> KEGG: " ,KEGG_Pathways))) +
  geom_point(alpha=0.7)+
  geom_label_repel(aes(label = KEGG_Pathways),
                  box.padding   = unit(0.35, "lines"), 
                  point.padding = unit(0.5, "lines"),
                  segment.color = 'grey50',
                  size=5.5) +
  scale_color_gradientn(colors = met.brewer("Gauguin",type = "continuous")) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "white"),
        panel.grid.minor = element_line(colour = "white"),
        axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left   = element_line(color = 'black'),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 15),
        axis.text=element_text(size=18),
        axis.title=element_text(size=18,face="bold"))+
  ylab("-log10(p.value)") +
  xlab("Pathway impact value")

ggsave2("Reprimidos_met.png",width = 8,height = 8)


Reprimidos_plotly_chlo <-ggplotly(Reprimidos_chlo,x = ~Impact, y = ~log10_p.value,
                             tooltip = "text",
                             width = 1100, height = 900)%>%
               layout(annotations = b)



# Static Bubbleplot
plot_grid(Activados_chlo,Reprimidos_chlo, nrow = 1, ncol = 2,align = "hv")
ggsave2("Enrichment.png",width = 25,height = 10)


# Interactive Bubbleplot

Enrich_interac <-subplot(Activados_plotly_chlo, Reprimidos_plotly_chlo ,titleX = T,titleY = T, margin = 0.1, nrows=1) 
htmlwidgets::saveWidget(as_widget(Enrich_interac), "Enrich_metabolite_interactive.html",selfcontained = T)


```

Barplots similar to clusterprofiler packages results were made.

```{r warning=F,echo=F}
# More relevatn pathways (more than 2 metabolites per pathway):

dim(pathway_results_act_chlo)
reduced_act_chlo <-pathway_results_act_chlo[pathway_results_act_chlo$Hits >= 2,]
dim(reduced_act_chlo)
head(reduced_act_chlo)

dim(pathway_results_rep_chlo)
reduced_rep_chlo<-pathway_results_rep_chlo[pathway_results_rep_chlo$Hits >= 2,]
dim(reduced_rep_chlo)
head(reduced_rep_chlo)

# Activados CHLO
Bar_Act_CHLO <- ggplot(reduced_act_chlo, aes(x=reorder(KEGG_Pathways,-log10_p.value), y=Hits, fill=log10_p.value)) + 
  geom_bar(stat = "identity") +
  # scale_fill_gradientn(colors = met.brewer("OKeeffe2",9))+
  scale_fill_continuous(low="red", high="blue",
                                  guide=guide_colorbar(reverse=TRUE))+
  coord_flip()+
  xlab(NULL) +
  ylab(NULL)+
  ggtitle("Activated Metabolite Enrichment")

# Reprimidos CHLO
Bar_Rep_CHLO <- ggplot(reduced_rep_chlo, aes(x=reorder(KEGG_Pathways,-log10_p.value), y=Hits, fill=log10_p.value)) + 
  geom_bar(stat = "identity") +
  # scale_fill_gradientn(colors = met.brewer("OKeeffe2",9))+
  scale_fill_continuous(low="red", high="blue",
                                  guide=guide_colorbar(reverse=TRUE))+
  coord_flip()+
  xlab(NULL) +
  ylab(NULL)+
  ggtitle("Repressed metabolite Enrichment")

# Estaticos Barplot
plot_grid(Bar_Act_CHLO,Bar_Rep_CHLO, nrow = 1, ncol = 2)
ggsave2("Barplots_Tartita.png",width = 15,height = 5)

```

